---
title: "assignment2_final"
author: "Melanie Nethercott"
date: "12/16/2017"
output: html_document
---

# Assignment 2 - Social Network Analysis

## Part I
Start by installing the "igraph" package. Once you have installed igraph, load the package.

Now upload the data file "discipline-data.csv" as a data frame called "D1". Each row is a disciplinary action from a teacher to a student so the first line shows that teacher "E" sent student "21" to the principal. It also shows the gender of both the teacher and student and the student's main elective field of study ("major"") and the field that the teacher instructs in ("t.expertise").

Before you proceed, you will need to change the data type of the student id variable. Since it is a number R will automatically think it is an integer and code it as such (look at the list of variables by clicking on the data frame arrow in the Data pane. Here you will see the letters "int"" next to the stid variable, that stands for integer). However, in this case we are treating the variable as a category, there is no numeric meaning in the variable. So we need to change the format to be a category, what R calls a "factor". We can do this with the following code:

## Part I
Installing relevant packages and uploading data. Change student id variable to be a factor rather than a numeric value.
```{r}
install.packages ("igraph")
library ("igraph")
library(dplyr)
library(tidyr)
D1 <- read.csv ("discipline-data.csv", sep = ",", header = TRUE) 
D1$stid <- as.factor(D1$stid)
```
Since our data represnts every time a teacher sends a student to the principal there are multiple rows when the same teacher sends the same student. We want to collapse these into a single row, with a variable that shows how many times a teacher-student pair appears.

```{r}
#Selecting key variables
D2 <- dplyr::select(D1, tid, stid)
#Collapsing rows to show total count for each teacher-student pair.
EDGE <- dplyr::count(D2, tid, stid)
#Setting names for EDGE data columns
names(EDGE) <- c("from", "to", "count")
```


```{r}
#Separating the teachers from our original data frame
V.TCH <- dplyr::select(D1, tid, t.gender, t.expertise)
#Removing all duplicate rows.
V.TCH <- unique(V.TCH)
#Adding a variable that describes that they are teachers
V.TCH$group <- "teacher"

#Repeating process for the students
V.STD <- dplyr::select(D1, stid, s.gender, s.major)
V.STD <- unique(V.STD)
V.STD$group <- "student"

#Make sure that the student and teacher data frames have the same variables names
names(V.TCH) <- c("id", "gender", "topic", "group")
names(V.STD) <- c("id", "gender", "topic", "group")

#Bind the two data frames together
VERTEX <- dplyr::bind_rows(V.TCH, V.STD)
```

```{r}
#Turning data into igraph objects. First we will make an object that contains the graph information using our two dataframes EDGE and VERTEX. 
g <- graph.data.frame(EDGE, directed=TRUE, vertices=VERTEX)

#Plot our graph.
plot(g,layout=layout.fruchterman.reingold)

#Color the nodes according to gender.
plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender)

#Change the thickness of the edge according to the number of times a particular teacher has sent a particular student to the principal.
plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender, edge.width=EDGE$n)
```

## Part II

In Part II your task is to [look up](http://igraph.org/r/) in the igraph documentation and create a graph that sizes the student vertices in terms of the number of disciplinary actions they have recieved, and the teachers in terms of the number of disciplinary actions they have given out.

```{r}
#Creating a new vertexsize table with total students given disciplinary action for each teacher and total received for each student and merging into VERTEX dataframe.
ttotal <- dplyr::count(D2, tid)
stotal <- dplyr::count(D2, stid)
names(ttotal) <- c("id", "total")
names(stotal) <- c("id", "total")
vertexsize <- bind_rows(ttotal, stotal)
VERTEX <- merge(vertexsize, VERTEX, by.x="total")

#Plotting the graph.
plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender, edge.width=EDGE$n, vertex.size=vertexsize$total)

#Plotting the graph to make smaller nodes more visible. Not sure this helps??
plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender, edge.width=EDGE$count, vertex.size=vertexsize$total*2, edge.arrow.size=0.4)
```

## Part III

Now practice with data from our class. Please create a **person-network** with the data set hudk4050-classes.csv. To create this network you will need to create a person-class matrix using the tidyr functions and then create a person-person matrix using `t()`. You will then need to plot a matrix rather than a data frame using igraph.

Once you have done this, also [look up](http://igraph.org/r/) how to generate the following network metrics: betweeness centrality and dregree. **Who is the most central person in the network?**

```{r}
#Combining first and last names in a new column, since some names are duplicated.
D3 <- read.csv ("hudk4050-classes.csv", sep = ",", header = TRUE) 
student<-paste(D3$First.name,"_",D3$Last.name)
D3 <- data.frame(student, D3)
D4 <- dplyr::select(D3, -Last.name, -First.name)

#Gathering all course columns into one and assigning 0 to missing values.
D4 <- gather(D4, key="student", value=2:6)
cols <- c("student","Course.No.", "Course.Name")
colnames(D4) <- cols
D4 <- arrange(D4, desc(Course.Name))
D4[D4==""] <- 0

#Abbreviating the names as the student names are long and may make the network messy.
names <- abbreviate(D4$student, minlength = 4, use.classes = TRUE,
           dot = FALSE, strict = FALSE,
           method = c("left.kept", "both.sides"), named = TRUE)
D4 <- data.frame(names, D4)
D4 <- dplyr::select(D4, -student)

#Removing HUDK4050 course (since everyone has that in common) and removing rows without Course.Name and Course.No values since these do not contain any information for plotting our network. 
D4 <- dplyr::select(D4, -Course.No.)
D4$Binary.Count <- ifelse(D4$Course.Name>0, 1, 0)
D4 <- subset(D4, Course.Name !="HUDK4050")
D4 <- unique(D4, incomprabales = "student")
D4 <- subset(D4, Binary.Count!="0")

#Using spread to have all the course names as columns and a 1 or 0 for each student depending on whether they took each class or not. 
D5 <- spread(D4, "Course.Name", "Binary.Count")
D5 [is.na(D5)] <- 0
row.names(D5)<-D5$names
D5 <- select(D5, -names)

#Turning the person-class dataframe into a person-class matrix.
m1 <- as.matrix(D5)

#Turning the person-class matrix into a person-person matrix. Turning diagonal to 0 since the diagonal just charts the classes shared between the same person. 
m2<-t(m1)
m3 <-m1 %*% m2
diag(m3)<-0
```


```{r}
#Plotting person to person network
scnetwork <- graph_from_adjacency_matrix (m3, mode = c("undirected"), weighted = TRUE, diag = FALSE, add.rownames = TRUE)
plot(scnetwork,layout=layout.fruchterman.reingold)

#Calculating betweeness centrality and degree. 
betweenness(scnetwork, v = V(scnetwork), directed = FALSE)
degree(scnetwork, v = V(scnetwork))
#Lauren, Rob, Jonathan, Zhuqian, and Josh have the highest betweeness. Chuheng, Dave, Devan, Jie, Jingtong, Jonathan, Josh, Lauren, Rob, Sam, Shreya and Zhuqian have the highest degree centrality. Since they have both the highest degree centrality and highest betweeness, Lauren, Rob, Jonathan, Josh and Zhuqian are the most central actors.
```



