---
title: "Assignment 2 Solutions"
output: html_document
---

Now upload the data file "discipline-data.csv" as a data frame called "D1". Each row is a disciplinary action from a teacher to a student so the first line shows that teacher "E" sent student "21" to the principal. It also shows the gender of both the teacher and student and the student's main elective field of study ("major"") and the field that the teacher instructs in ("t.expertise").



```{r}
rm(list=ls())
D1=read.csv("discipline-data.csv",header = T)
attach(D1)
D1$stid <- as.factor(D1$stid)

```

convert our data into an edge list!
First we will isolate the variables that are of interest: tid and stid

```{r}
library(dplyr)

D2 <- dplyr::select(D1, tid, stid)
```

Since our data represnts every time a teacher sends a student to the principal there are multiple rows when the same teacher sends the same student. We want to collapse these into a single row, with a variable that shows how many times a teacher-student pair appears.

```{r}
EDGE <- dplyr::count(D2, tid, stid)

names(EDGE) <- c("from", "to", "count")
```

EDGE is your edge list. Now we need to make the vertex list, a list of all the teachers and students and their characteristics in our network.

```{r}
#First we will separate the teachers from our original data frame
V.TCH <- dplyr::select(D1, tid, t.gender, t.expertise)
#Remove all the repeats so that we just have a list of each teacher and their characteristics
V.TCH <- unique(V.TCH)
#Add a variable that describes that they are teachers
V.TCH$group <- "teacher"

#Now repeat this process for the students
V.STD <- dplyr::select(D1, stid, s.gender, s.major)
V.STD <- unique(V.STD)
V.STD$group <- "student"

#Make sure that the student and teacher data frames have the same variables names
names(V.TCH) <- c("id", "gender", "topic", "group")
names(V.STD) <- c("id", "gender", "topic", "group")

#Bind the two data frames together (you will get a warning because the teacher data frame has 5 types of id (A,B,C,D,E) and the student has 25 (1-30), this isn't a problem)
VERTEX <- dplyr::bind_rows(V.TCH, V.STD)
```
Now we have both a Vertex and Edge list it is time to plot our graph!

```{r}
#Load the igraph package

library(igraph)

#First we will make an object that contains the graph information using our two dataframes EDGE and VERTEX. Notice that we have made "directed = TRUE" - our graph is directed since discipline is being given from a teacher to a student.

g <- graph.data.frame(EDGE, directed=TRUE, vertices=VERTEX)

#Now we can plot our graph using the force directed graphing technique - our old friend Fruchertman-Reingold!

plot(g,layout=layout.fruchterman.reingold)

#There are many ways to change the attributes of the graph to represent different characteristics of the newtork. For example, we can color the nodes according to gender.

plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender)

#We can change the thickness of the edge according to the number of times a particular teacher has sent a particular student to the principal.

plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender, edge.width=EDGE$count)
```
## Part II

In Part II your task is to [look up](http://igraph.org/r/) in the igraph documentation and create a graph that sizes the student vertices in terms of the number of disciplinary actions they have recieved, and the teachers in terms of the number of disciplinary actions they have given out.



```{r}
# student and the number of disciplinary actions 
library(tidyr)

STDtotal=group_by(count(D2, stid))

STDtotal

TCHtotal=group_by(count(D2,tid))

TCHtotal

names(TCHtotal)[1]=c("ID")
names(STDtotal)[1]=c("ID")

Total=rbind(STDtotal,TCHtotal)
head(Total)



g1 <- graph.data.frame(EDGE, directed=TRUE, vertices=Total)


plot(g1,layout=layout.fruchterman.reingold, vertex.size=Total$n*2)


```

## Part III

Now practice with data from our class. Please create a **person-network** with the data set hudk4050-classes.csv. To create this network you will need to create a person-class matrix using the tidyr functions and then create a person-person matrix using `t()`. You will then need to plot a matrix rather than a data frame using igraph.



```{r}
rm(list=ls())
data=read.csv("HUDK4050-2018-classes.csv")
ncol(data)

library(tidyr)
library(dplyr)
mydata1=select(data,15:21)
mydata2=unite(mydata1,"ID",c("Q8","Q9"))
mydata=mydata2[-c(1,2,3,39), ]
names(mydata)=c("ID","Class 1","Class 2", "Class 3","Class 4","Class 5")


d1=gather(mydata,class.number, class,"Class 1","Class 2", "Class 3","Class 4","Class 5",na.rm=T,convert = F)

d2=select(d1,ID,class)

# we take out the HUDK 4050, because everyone has it
d3=filter(d2,class > 0,class!="HUDK 4050")

# now we want to count the people who take the same course

d3$count<-1

d4=spread(d3,class,count)

# now we need a matrix 
row.names(d4)<-d4$ID

d4$ID<-NULL

d5=ifelse(is.na(d4),0,1)

d6=as.matrix(d5)

d7=d6%*%t(d6)


diag(d7)=NA

g=graph.adjacency(d7,mode="undirected")

library(ggplot2)

g1=plot(g,layout=layout.fruchterman.reingold,vertex.size=2)


degree(g)


betweenness(g)



```

