# Assignment 2 - Social Network Analysis

## Part I
Start by installing the "igraph" package. Once you have installed igraph, load the package.

Now upload the data file "discipline-data.csv" as a data frame called "D1". 
```{r}
discipline.data <- read.csv("discipline-data.csv")
D1 <- discipline.data
```

Each row is a disciplinary action from a teacher to a student so the first line shows that teacher "E" sent student "21" to the principal. It also shows the gender of both the teacher and student and the student's main elective field of study ("major"") and the field that the teacher instructs in ("t.expertise").

Before you proceed, you will need to change the data type of the student id variable. Since it is a number R will automatically think it is an integer and code it as such (look at the list of variables by clicking on the data frame arrow in the Data pane. Here you will see the letters "int"" next to the stid variable, that stands for integer). However, in this case we are treating the variable as a category, there is no numeric meaning in the variable. So we need to change the format to be a category, what R calls a "factor". We can do this with the following code:

```{r}
D1$stid <- as.factor(D1$stid)
```

igraph requires data to be in a particular structure. There are several structures that it can use but we will be using a combination of an "edge list" and a "vertex list". As you might imagine the edge list contains a list of all the relationships between students and teachers and any characteristics of those edges that we might be interested in. There are two essential variables in the edge list a "from" variable and a "to" variable that descibe the relationships between vertices (a disciplinary action is given "from" and teacher "to" a student). While the vertix list contains all the characteristics of those vertices, in our case gender and major.

So let's convert our data into an edge list!

First we will isolate the variables that are of interest: tid and stid

```{r}
library(dplyr)

D2 <- select(D1, tid, stid)
```

Since our data represnts every time a teacher sends a student to the principal there are multiple rows when the same teacher sends the same student. We want to collapse these into a single row, with a variable that shows how many times a teacher-student pair appears.

```{r}

EDGE <- count(D2, tid, stid)

names(EDGE) <- c("from", "to", "count")

```

EDGE is your edge list. Now we need to make the vertex list, a list of all the teachers and students and their characteristics in our network.

```{r}

#First we will separate the teachers from our original data frame
V.TCH <- select(D1, tid, t.gender, t.expertise)
#Remove all the repeats so that we just have a list of each teacher and their characteristics
V.TCH <- unique(V.TCH)
#Add a variable that describes that they are teachers
V.TCH$group <- "teacher"

#Now repeat this process for the students
V.STD <- select(D1, stid, s.gender, s.major)
V.STD <- unique(V.STD)
V.STD$group <- "student"

#Make sure that the student and teacher data frames have the same variables names
names(V.TCH) <- c("id", "gender", "topic", "group")
names(V.STD) <- c("id", "gender", "topic", "group")

#Bind the two data frames together (you will get a warning because the teacher data frame has 5 types of id (A,B,C,D,E) and the student has 25 (1-30), this isn't a problem)
VERTEX <- bind_rows(V.TCH, V.STD)

```

Now we have both a Vertex and Edge list it is time to plot our graph!

```{r}
#Load the igraph package

library(igraph)

#First we will make an object that contains the graph information using our two dataframes EDGE and VERTEX. Notice that we have made "directed = TRUE" - our graph is directed since discipline is being given from a teacher to a student.

g <- graph.data.frame(EDGE, directed=TRUE, vertices=VERTEX)

#Now we can plot our graph using the force directed graphing technique - our old friend Fruchertman-Reingold!

plot(g,layout=layout.fruchterman.reingold)

#There are many ways to change the attributes of the graph to represent different characteristics of the newtork. For example, we can color the nodes according to gender.

plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender)

#We can change the thickness of the edge according to the number of times a particular teacher has sent a particular student to the principal.

plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender, edge.width=EDGE$count)

````

## Part II

In Part II your task is to [look up](http://igraph.org/r/) in the igraph documentation and create a graph that sizes the student vertices in terms of the number of disciplinary actions they have recieved, and the teachers in terms of the number of disciplinary actions they have given out.

---------------------

Ok real talk all of my original code got deleted at some point? but for whatever reason I had saved the environment SO workin' it backwards. 

How did VERTEX become VERTEX_3?
well I think VERTEX_2 was actually nonsense... a stepping stone on my way to clarity
I made VERTEX_S for students with a column of disciplinary actions
then I made VERTEX_T for teachers with a column of disciplinary actions
Then I remember giggling to myself making VERTEX_2, which I don't even think I ended up using for anything relevant because the real deal is VERTEX_3, which was a merged df of VERTEX_S AND VERTEX_T

``````{r}
edge2T <- count(D2, tid)
names(edge2T) <- c("ID", "Count")
edge2S <- count(D2, stid)
names(edge2S) <- c("ID", "Count")
edge_test <- bind_rows(edge2T, edge2S)

names(VERTEX) <- c("ID", "gender", "topic", "group")

VERTEX_2 <- merge(VERTEX, edge_test,by="ID")

VERTEX_2 <- mutate(VERTEX_2, Count_Graph_Size = Count + 14)

g2 <- graph.data.frame(EDGE, directed=TRUE, vertices=VERTEX_2)

plot(g2, layout = layout_with_gem, vertex.size = VERTEX_2$Count_Graph_Size, vertex.color = (ifelse(test=VERTEX_2$group=="student", yes="yellow", no="orange")), edge.arrow.size=0.3, vertex.label.cex=0.6, vertex.label.family="sans")
```


## Part III

Now practice with data from our class. Please create a **person-network** with the data set hudk4050-classes.csv. To create this network you will need to create a person-class matrix using the tidyr functions and then create a person-person matrix using `t()`. You will then need to plot a matrix rather than a data frame using igraph.

select necessary columns, remove first 3 rows of nonsense

```{r}
df <- hudk4050.classes
df <- select(df, V20:V26)
df <- df[-c(1:3), ]
```

gather class columns (2:7) by number
rename headers
remove number
add Count column

```{r}
df1 <- gather(df, "number", "Course", 2:7)
names(df1) <- c("UNI", "remove", "Course")
df1 <- select(df1, UNI, Course)
df1$Count = 1
```

CLEAN IT UP, KID 

remove space or other char. between Courses XXXX 1234
remove rows with blank Course data 
remove rows with HUDK4050
update course names
remove student who didn't put any letters w/course #s 

```{r}
df1$Course = gsub(" ", "", df1$Course)
df1$Course = gsub("-", "", df1$Course)
df1 <- filter(df1, Course !="")
df1 <- filter(df1, Course !="HUDK4050")

clean <- mutate(df1, nchar = nchar(df1$Course))
clean <- arrange(clean, desc(nchar))
count(clean, nchar)
filter(clean, nchar !=8)


df1$Course[df1$Course=="G5072"] <- "GR5072"
df1$Course[df1$Course=="G5067"] <- "GR5067"
df1$Course[df1$Course=="QMSS5067"] <- "GR5072"
df1$Course[df1$Course=="QMSSGR5067"] <- "GR5067"
df1$Course[df1$Course=="QMSSG5015"] <- "GR5015"
df1$Course[df1$Course=="IFSF4090002"] <- "ITSF4090"
df1$Course[df1$Course=="QMSS5010"] <- "GR5010"
df1$Course[df1$Course=="QMSS5015"] <- "GR5015"
df1$Course[df1$Course=="QMSS5021"] <- "GR5021"
df1$Course[df1$Course=="QMSS5072"] <- "GR5072"
df1$Course[df1$Course=="QMSSG5072"] <- "GR5072"

df1 <- filter(df1, UNI !="ZC2505")
df1 <- arrange(df1, Course)


course_count <- count(df1, Course)
```

# spread df1 into m 
rename rows as UNIs 
remove UNI column...
  Q: is doing X1$UNI = NULL the same as doing X2 <- select(X1, -UNI)?
change NAs to 0s 

create m1 as multiple of m by its inverse
change diag of m1 to NA

create g_m1 with graph.adjacency function (undirected)
  adding remove loops cut down edges to 438 from 646... 
  g_m1 <- simplify(g_Xm, remove.multiple = T, remove.loops = T)

```{r}
m <- spread(df1, Course, Count)
row.names(m) <- m$UNI
m$UNI = NULL
m <- ifelse(is.na(m), 0, 1)

m1 <- as.matrix(m) %*% t(as.matrix(m))
diag(m1) <- NA

g_m1 <- graph.adjacency(m1, mode = "undirected")

E(g_m1)
g_m2 <- simplify(g_m1, remove.multiple = T, remove.loops = T)
E(g_m2)
```

plot it up. don't love fruchterman reingold for this. Tried a bunch, fav is kk despite this layout hammering home all of my loneliness insecurities. #nobodyputsbabyinacorner
Also, UNIs are too long here. should have assigned shorter IDs, ESP for full email kid.

```{r}
plot(g_m2, layout=layout_with_kk, edge.width = 0.25, vertex.label.cex=0.35, vertex.label.family="sans", vertex.size = 8, margin = -.05, asp = -.5)
```

Once you have done this, also [look up](http://igraph.org/r/) how to generate the following network metrics: betweeness centrality and degree. **Who is the most central person in the network?**

So these results for degree are kinda unsettling because I copied and pasted my code with minor alterations from a different Rmd file i had going and the numbers are a bit off... 
Here's what I had written for the other file's data, where student ll3358 had a degree of 48: 
ll3358 is taking 3 courses outside of HUDK4050, which are the 1st, 2nd, and 5th most popular classes being taken by HUDK4050 students this semester. HUDM5026 (18 students),	HUDK4052 (21 students),	HUDK4029 (12 students). 
```{r}
sort(degree(g_m2))
```



Betweenness is more or less the same as my other markdown doc, so maybe this one is correct and the other one was wacked. Unclear, anywho:

yz3679 is taking 4 courses outside of HUDK4050:	HUDK4052,	HUDK4080,	HUDM5026, and	MSTU4052. HUDK4052 and HUDM5026 are the top 2 most popular classes shared among HUDK4050 students, then being in HUDM4080 connects yz3679 to 2 other students (ql2360 and IB2445) who are both in HUDK4029 (the 5th most popular class), plus ql2360 is in the 6th most popular class, and IB2445 in another random class that ties back to one other student who is fairly well connected as well (zz2726, betweenness = 13.1). I don't know, sounds about right though. 


```{r}
sort(betweenness(g_m2))
```

the most central person in the network is yz3679. don't tell your secrets to yz3679.


### To Submit Your Assignment

Please submit your assignment by first "knitting" your RMarkdown document into an html file and then comit, push and pull request both the RMarkdown file and the html file.