# Assignment 2 - Social Network Analysis

## Part I
Start by installing the "igraph" package. Once you have installed igraph, load the package.
```{r}
#install.packages("igraph")
library(igraph)
```

Now upload the data file "discipline-data.csv" as a data frame called "D1". Each row is a disciplinary action from a teacher to a student so the first line shows that teacher "E" sent student "21" to the principal. It also shows the gender of both the teacher and student and the student's main elective field of study ("major"") and the field that the teacher instructs in ("t.expertise").

```{r}
#name the given data file
D1 <- read.csv("discipline-data.csv")
```


Before you proceed, you will need to change the data type of the student id variable. Since it is a number R will automatically think it is an integer and code it as such (look at the list of variables by clicking on the data frame arrow in the Data pane. Here you will see the letters "int"" next to the stid variable, that stands for integer). However, in this case we are treating the variable as a category, there is no numeric meaning in the variable. So we need to change the format to be a category, what R calls a "factor". We can do this with the following code:

```{r}
D1$stid <- as.factor(D1$stid)
```

igraph requires data to be in a particular structure. There are several structures that it can use but we will be using a combination of an "edge list" and a "vertex list". As you might imagine the edge list contains a list of all the relationships between students and teachers and any characteristics of those edges that we might be interested in. There are two essential variables in the edge list a "from" variable and a "to" variable that descibe the relationships between vertices (a disciplinary action is given "from" and teacher "to" a student). While the vertix list contains all the characteristics of those vertices, in our case gender and major.

So let's convert our data into an edge list!

First we will isolate the variables that are of interest: tid and stid

```{r}
library(dplyr)
#get the D2 file
D2 <- select(D1, tid, stid)
```

Since our data represnts every time a teacher sends a student to the principal there are multiple rows when the same teacher sends the same student. We want to collapse these into a single row, with a variable that shows how many times a teacher-student pair appears.

```{r}
EDGE <- count(D2, tid, stid)
names(EDGE) <- c("from", "to", "count")
```

EDGE is your edge list. Now we need to make the vertex list, a list of all the teachers and students and their characteristics in our network.

```{r}
#select teachers data and separate from original sheet
V.TCH <- select(D1, tid, t.gender, t.expertise)
#get a clean list by delete all repeats data
V.TCH <- unique(V.TCH)
#makes the group shows as teachers
V.TCH$group <- "teacher"

#follow the steps what teachers did
V.STD <- select(D1, stid, s.gender, s.major)
V.STD <- unique(V.STD)
V.STD$group <- "student"

#check same variables names for both teachers and students
names(V.TCH) <- c("id", "gender", "topic", "group")
names(V.STD) <- c("id", "gender", "topic", "group")

#bind teahcers and students' data together
VERTEX <- bind_rows(V.TCH, V.STD)

```

Now we have both a Vertex and Edge list it is time to plot our graph!

```{r}
#install igraph package

library(igraph)

#teachers to studnents always have directions，so we can get a graph include all informations

g <- graph.data.frame(EDGE, directed=TRUE, vertices=VERTEX)

plot(g,layout=layout.fruchterman.reingold)

#Now we have to color the nodes based on gender.

gender_colours =ifelse(VERTEX$gender == "female", yes="blue",no = "red")

VERTEX = cbind(VERTEX,gender_colours)

plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender_colours)

plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender_colours, edge.width=EDGE$count)


````

## Part II

In Part II your task is to [look up](http://igraph.org/r/) in the igraph documentation and create a graph that sizes the student vertices in terms of the number of disciplinary actions they have recieved, and the teachers in terms of the number of disciplinary actions they have given out.

```{r}

#when we get teachers and students total number, it helps we count

s_total<- EDGE %>% group_by(to) %>% summarise(sum(count))
names(s_total)<-c("id","count")
t_total<-EDGE %>% group_by(from) %>% summarise(sum(count))
names(t_total)<-c("id","count")

VERTEX_new<-bind_rows(s_total,t_total)
VERTEX_new

library(igraph)

#graph it
g<- graph.data.frame(EDGE,directed = TRUE,vertices = VERTEX_new)

#plot everything

plot(g,layout = layout.fruchterman.reingold,
     edge.width = EDGE$count,
      vertex.size = VERTEX_new$count*1.5)

```


## Part III

Now practice with data from our class. Please create a **person-network** with the data set hudk4050-classes.csv. To create this network you will need to create a person-class matrix using the tidyr functions and then create a person-person matrix using `t()`. You will then need to plot a matrix rather than a data frame using igraph.
```{r}

library(igraph)
library(dplyr)
library(tidyr)

#name the file database

database<-read.csv("hudk4050-classes.csv",header = TRUE)

#variable selection

c_type<- database %>% select(First.Name = 'First.Name',
                                Last.Name ='Last.Name',
                                class1 = 'Class.1',
                                class2 = 'Class.2',
                                class3 = 'Class.3',
                                class4 = 'Class.4',
                                class5 = 'Class.5',
                                class6 = 'Class.6'
                                )

c_type<-as.data.frame(c_type)

c_type<-unite(c_type,Name,"First.Name","Last.Name",sep="")

#filter the data

#I changed Zimo's data with the original title to fit

gathering<-gather(c_type,class.number,class.Name,-Name)

gathering<-filter(gathering,class.Name>0)

gathering<-select(gathering,Name,class.Name)

gathering1111<-function(a){gsub("","",a)}

gathering<-lapply(gathering,gathering1111)

gathering$count<-1

gathering<-as.data.frame(gathering)

gathering<-filter(gathering,class.Name!="HUDK4050")

gathering$class.Name<- gsub(pattern = "QMSS-",
                                replacement = "",  
                                x=gathering$class.Name)

gathering$class.Name <- gsub(pattern = "QMSS",
                                     replacement = "G",
                                     x = gathering$class.Name)
gathering$class.Name <- gsub(pattern = "GG",
                                     replacement = "G",
                                     x = gathering$class.Name)
gathering$class.Name <- gsub(pattern = "GR",
                                     replacement = "G",
                                     x = gathering$class.Name)

#make output unique and make matrix

gathering<-unique(gathering)

g_matrix<- spread(gathering,class.Name,count,fill=0)

#give it a new name

row.names(g_matrix)<-g_matrix$Name

g_matrix$Name<-NULL

g_matrix<-as.matrix(g_matrix)

g_matrix2<-g_matrix %*% t(g_matrix)

diag(x=g_matrix2)<- NA

chart_d<- graph.adjacency(g_matrix2,mode = "undirected")

#Plog everything

plot.igraph(chart_d, layout = layout.fruchterman.reingold, vertex.size = 12,
            vertex.label.cex = 0.4)

```

Once you have done this, also [look up](http://igraph.org/r/) how to generate the following network metrics: betweeness centrality and dregree. **Who is the most central person in the network?**

```{r}

betweeness_centrality<- betweenness(chart_d)
betweeness_centrality[betweeness_centrality == max(betweeness_centrality)]

degrees <- degree(chart_d)
degrees[degrees == max(degrees)]
```

### To Submit Your Assignment

Please submit your assignment by first "knitting" your RMarkdown document into an html file and then comit, push and pull request both the RMarkdown file and the html file.
