# Assignment 2 - Social Network Analysis

title: "Assignment 2"
author: "Abdul Abad"
date: "December, 18 2018"
output: html_document
## Part I
Start by installing the "igraph" package. Once you have installed igraph, load the package.


```{r}
library(igraph)
```


Now upload the data file "discipline-data.csv" as a data frame called "D1". Each row is a disciplinary action from a teacher to a student so the first line shows that teacher "E" sent student "21" to the principal. It also shows the gender of both the teacher and student and the student's main elective field of study ("major"") and the field that the teacher instructs in ("t.expertise").

```{r}
D1 <- read.csv("discipline-data.csv")
```

Before you proceed, you will need to change the data type of the student id variable. Since it is a number R will automatically think it is an integer and code it as such (look at the list of variables by clicking on the data frame arrow in the Data pane. Here you will see the letters "int"" next to the stid variable, that stands for integer). However, in this case we are treating the variable as a category, there is no numeric meaning in the variable. So we need to change the format to be a category, what R calls a "factor". We can do this with the following code:

```{r}
@@ -80,15 +94,80 @@ plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender)
#We can change the thickness of the edge according to the number of times a particular teacher has sent a particular student to the principal.
plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender, edge.width=EDGE$count)
````
```

## Part II

In Part II your task is to [look up](http://igraph.org/r/) in the igraph documentation and create a graph that sizes the student vertices in terms of the number of disciplinary actions they have recieved, and the teachers in terms of the number of disciplinary actions they have given out.


```{r}
#disciplinary actions
teachers <- dplyr::count(D2, tid)
students <- dplyr::count(D2, stid) 
colnames(teachers) <- colnames(students) <- c("id","size")
VERTEX_sizes <- dplyr::bind_rows(teachers, students)
VERTEX_joined <-  left_join(VERTEX, VERTEX_sizes)
g <- graph.data.frame(EDGE, directed=TRUE, vertices=VERTEX_joined)
#disciplinary actions for size
plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX_joined$gender, edge.width=EDGE$count, size = VERTEX_joined$size )
```


## Part III

Now practice with data from our class. Please create a **person-network** with the data set hudk4050-classes.csv. To create this network you will need to create a person-class matrix using the tidyr functions and then create a person-person matrix using `t()`. You will then need to plot a matrix rather than a data frame using igraph.

Once you have done this, also [look up](http://igraph.org/r/) how to generate the following network metrics: betweeness centrality and dregree. **Who is the most central person in the network?**


```{r}
library(tidyr)
#load data
data <- read.csv("HUDK4050-2018-classes.csv", skip = 3)
headers <- read.csv("HUDK4050-2018-classes.csv", skip = 1)
#set column names
headers <- colnames(headers)
colnames(data) <- headers
#subset pearson and class
data <- mutate(data, person = paste(First.Name, Last.Name))
D <-  select(data, Class.1:Class.5,person)
d1 <- gather(D, class, name, -person)
#data cleaning
d1 <- filter(d1,  name != "" & person !=" ")
#for examples: HUDK 4050 and HUDK4050 are the same
d1 <- mutate(d1, name = gsub(" ","",name)) 
d2 <- select(d1, person, name)
d2 <- mutate(d2,  has = 1)
d2 <- unique(d2)
full <-  expand(d2, person, name)
#merged
d3 <- left_join(full, d2)
d3 <- mutate(d3, has = ifelse(is.na(has),0,has))
#person-class matrix
d4 <- spread(d3, name, has)
mat <- as.matrix(d4[,-1])
row.names(mat) <- data.frame(d4)[,1]
colnames(mat) <- colnames(d4[,-1])
#person-person matrix
mat2 <- mat %*% t(mat)
diag(mat2) <- 0
#graph from person-person matrix
g <- graph_from_adjacency_matrix(mat2, mode = "undirected", weighted = TRUE)
plot(g,layout=layout.fruchterman.reingold)
betweenness(g)
degree(g)
which.max(betweenness(g))
```

The person is Delciane Wright has the largest betweeness centrality.