---
title: "Assignment 2 - Social Network Analysis"
author: "Charles Lang"
date: "Oct 16, 2016"
output: html_document
---

In Assignment 2 we will be looking at some disciplinary data from a middle school. The file "discipline-data.csv" shows which teacher sent which student to the principal during the semester.

We will be using the "igraph" package to visualize the disciplinary relationships between teachers and students as a network. You can read more about igraph here:

http://igraph.org/r/

Start by installing the "igraph" package. Once you have installed igraph, load the package.

```{r}
#Clears memory
rm(list = ls())

#Clear the console
cat("\014")

#install.packages("igraph")
library("igraph")
```

Now upload the data file "discipline-data.csv" as a data frame called "D1". Each row is a disciplinary action from a teacher to a student so the first line shows that teacher "E" sent student "21" to the principal. It also shows the gender of both the teacher and student and the student's main elective field of study ("major"") and the field that the teacher instructs in ("t.expertise").

```{r}
D1<-read.csv("discipline-data.csv")
names(D1)
```

Before you proceed, you will need to change the data type of the student id variable. Since it is a number R will automatically think it is an integer and code it as such (look at the list of variables by clicking on the data frame arrow in the Data pane. Here you will see the letters "int"" next to the stid variable, that stands for integer). However, in this case we are treating the variable as a category, there is no numeric meaning in the variable. So we need to change the format to be a category, what R calls a "factor". We can do this with the following code:

```{r}
D1$stid <- as.factor(D1$stid)
```

igraph requires data to be in a particular structure. There are several structures that it can use but we will be using a combination of an "edge list" and a "vertex list". As you might imagine the edge list contains a list of all the relationships between students and teachers and any characteristics of those edges that we might be interested in. There are two essential variables in the edge list a "from" variable and a "to" variable that descibe the relationships between vertices (a disciplinary action is given "from" and teacher "to" a student). While the vertix list contains all the characteristics of those vertices, in our case gender and major.

So let's convert our data into an edge list!

First we will isolate the variables that are of interest: tid and stid

```{r}
library(dplyr)

D2 <- dplyr::select(D1, tid, stid)
```

Since our data represnts every time a teacher sends a student to the principal there are multiple rows when the same teacher sends the same student. We want to collapse these into a single row, with a variable that shows how many times a teacher-student pair appears.

```{r}

EDGE <- dplyr::count(D2, tid, stid)

names(EDGE) <- c("from", "to", "count")

```

EDGE is your edge list. Now we need to make the vertex list, a list of all the teachers and students and their characteristics in our network.

```{r}

#First we will separate the teachers from our original data frame
V.TCH <- dplyr::select(D1, tid, t.gender, t.expertise)
#Remove all the repeats so that we just have a list of each teacher and their characteristics
V.TCH <- unique(V.TCH)
#Add a variable that describes that they are teachers
V.TCH$group <- "teacher"

#Now repeat this process for the students
V.STD <- dplyr::select(D1, stid, s.gender, s.major)
V.STD <- unique(V.STD)
V.STD$group <- "student"

#Make sure that the student and teacher data frames have the same variables names
names(V.TCH) <- c("id", "gender", "topic", "group")
names(V.STD) <- c("id", "gender", "topic", "group")

#Bind the two data frames together (you will get a warning because the teacher data frame has 5 types of id (A,B,C,D,E) and the student has 25 (1-30), this isn't a problem)
VERTEX <- dplyr::bind_rows(V.TCH, V.STD)

```

Now we have both a Vertex and Edge list it is time to plot our graph!

```{r}
#Load the igraph package

#library(igraph)

#First we will make an object that contains the graph information using our two dataframes EDGE and VERTEX. Notice that we have made "directed = TRUE" - our graph is directed since discipline is being given from a teacher to a student.

g <- graph.data.frame(EDGE, directed=TRUE, vertices=VERTEX)

#Now we can plot our graph using the force directed graphing technique - our old friend Fruchertman-Reingold!

plot(g,layout=layout.fruchterman.reingold)

#There are many ways to change the attributes of the graph to represent different characteristics of the newtork. For example, we can color the nodes according to gender.

plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender)

#We can change the thickness of the edge according to the number of times a particular teacher has sent a particular student to the principal.

plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender, edge.width=EDGE$count)

#Other attributes you may want to change are:

#Arrow size
plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender, edge.width=EDGE$count,edge.arrow.size=0.5)

#Remove the labels
plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender, edge.width=EDGE$count,edge.arrow.size=0.5,vertex.label=NA)

#Vertex size
plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender, edge.width=EDGE$count,edge.arrow.size=0.5,vertex.label=NA,vertex.size=10)
  
#And many others!
```

Play around with different graphs and then save your final version to the repo. Create a Zotero item for this R program and then commit, push and pull.

Once you have done this you need to practice with data from our class. Please create a **person-network** with the data set hudk4050-classes.csv. Once you have done this, also [look up](http://igraph.org/r/) how to generate the following network metrics: betweeness centrality and dregree. **Who is the most central person in the network?**

```{r}
#Load the data for HUDK-4050 class

D4050<-read.csv("hudk4050-classes.csv")
names(D4050)

#We transform it to characters first so we don't have issues with the different levels (will transform them back to factors later)
D4050$Course1<-as.character(D4050$Course1)
D4050$Course2<-as.character(D4050$Course2)
D4050$Course3<-as.character(D4050$Course3)
D4050$Course4<-as.character(D4050$Course4)
D4050$Course5<-as.character(D4050$Course5)

#We will use the last names of the students as their IDs + their first letter of their name (I'll change my last name to my first last name only to make it more pretty):
D4050$id<-as.character(D4050$Last.name)
D4050$id[D4050$id=="Bennett Colomer"]<-"Bennett"
D4050$id[D4050$Last.name=="Zhang" & D4050$First.name=="Jiaqing"]<-"Zhang_J"
D4050$id[D4050$Last.name=="Zhang" & D4050$First.name=="Xiaoyue"]<-"Zhang_X"

D4050$id<-as.factor(D4050$id)

#Some people included the course HUDK4050 and others did not, so I'll add it to their course list if they didn't:
for(i in 1:nrow(D4050)){
  #Retrieve the courses for each students
  courses=D4050[i,-c(1,2)]
  #Find the course "HUDK4050" within their courses
  aux=which(courses=="HUDK4050")
  
  #If they did not add HUDK4050, I'll add it in the first blank space they have
  if(length(aux)==0){
    D4050[i,which(courses=="")[1]+2]="HUDK4050"
  }
}

#Now, we subset are dataframe using the variables we need.
D4050_s<-dplyr::select(D4050,id,Course1,Course2,Course3,Course4,Course5)

#We create the edges, that in this case will be whether people belong to the same class (and how many classes do they belong to together)
D4050_r<-tidyr::gather(D4050_s,key=N_course,value=course,-id,factor_key=TRUE)
names(D4050_r)<-c("id","N_course","course")

#Drop the courses that are empty
D4050_r<-D4050_r[D4050_r$course!="",]

#Transform courses back to a factor
D4050_r$course<-as.factor(D4050_r$course)

#id_course<-as.data.frame(table(D4050_r$id,D4050_r$course))

#Now we create a matrix of edges (undirected):
#I'm sure that there is a more efficient way to do this, uut I cannot think of one right now.
SN_mat<-matrix(0,nrow=length(unique(D4050_r$id)),ncol=length(unique(D4050_r$id)))

for(i in 1:(length(unique(D4050$id))-1)){
  for(j in (i+1):length(unique(D4050$id))){
    aux=D4050_r[D4050_r$id==unique(D4050$id)[i] | D4050_r$id==unique(D4050$id)[j],]
    SN_mat[i,j]=sum(dplyr::count(aux,course)[,2]-1)
    
    #SN_mat[j,i]=SN_mat[i,j]
  }
}

#SN will now store the connections between students depending of the number of courses in which they are togheter
SN<-as.data.frame(t(SN_mat))
SN<-cbind(levels(D4050_r$id),SN)
names(SN)<-c("id",levels(D4050_r$id))

#Now we get a dataframe similar to EDGES (and we drop the node-to-same-node connections, and repeated connections)
edge4050 <- tidyr::gather(SN,st2,count,-id)
edge4050<-edge4050[edge4050$id!=edge4050$st2,]
edge4050<-edge4050[edge4050$count!=0,]

#Because this network is non-directional, we include both to-from combinations.
names(edge4050) <- c("from", "to", "count")


#Now, we wil create the vertex dataframe

vertex4050<-levels(D4050_s$id)
#Calculate the number of courses each student is taking
n_courses<-rowSums(D4050_s[,c("Course1","Course2","Course3","Course4","Course5")]!="")

vertex4050<-as.data.frame(cbind(vertex4050,n_courses))

names(vertex4050)<-c("id","n_courses")

vertex4050$id<-as.factor(vertex4050$id)

```
Finally, we can start *plotting!*

```{r}

g <- graph.data.frame(edge4050, directed=FALSE, vertices=vertex4050)

plot(g,layout=layout.fruchterman.reingold)
plot(g,layout=layout.fruchterman.reingold,vertex.color=vertex4050$n_courses)

#We will exclude the connection given by HUDK4050:

edge4050_2<-edge4050
edge4050_2$count<-edge4050$count-1
edge4050_2<-edge4050_2[edge4050_2$count!=0,]

g2 <- graph.data.frame(edge4050_2, directed=FALSE, vertices=vertex4050)

colrs <- adjustcolor( c("gray50", "tomato", "gold", "yellowgreen","purple"), alpha=.6)


#Graph showing the connections between the students (besides HUDK4050).

plot(g2,layout=layout.fruchterman.reingold, vertex.color=colrs[vertex4050$n_courses],edge.width=edge4050_2$count,vertex.size=5,xlim=c(-0.8,0.8),ylim=c(-0.7,0.7),vertex.label.cex = 0.8,asp=0.35,margin=0.1)
legend('bottomleft',legend=levels(vertex4050$n_courses),col=colrs,fill=colrs,bty="n",cex=0.5,horiz=F,title="N courses taken")

```

It is clear to see from the previous graph, that students that are taking more classes are more likely to be connected. Also, there is a clear cluster of students who are taking the same classes.

Now, we will calculate the betweenness centrality and degree of the network:

```{r}
#"The vertex and edge betweenness are (roughly) defined by the number of geodesics (shortest paths) going through a vertex or an edge"
estimate_betweenness(g2,directed=FALSE,cutoff=0,weights=edge4050_2$count)

#The last graph did not consider that all the nodes were connected, so we use the firs graph which includes the connection from HUDK4050
estimate_betweenness(g,directed=FALSE,cutoff=0,weights=edge4050$count)

#We can also calculate the closeness of the network, where closeness is defined as: "Cloness centrality measures how many steps is required to access every other vertex from a given vertex."
closeness(g,weights = edge4050$count, normalized = T)

names(closeness(g,weights = edge4050$count, normalized = T))[closeness(g,weights = edge4050$count, normalized = T)==min(closeness(g,weights = edge4050$count, normalized = T))]

#Here we can see that the people within the cluster, have the same value of closeness centrality.

#Finally, the degree of the network:
degree(g, loops = TRUE, normalized = TRUE)
#For the first graph, the degree is 1, because all students are directly connected through HUDK4050.

#Now, if we ignore the connection fot he HUDK4050 course:

degree(g2, loops = TRUE, normalized = TRUE)

names(degree(g2, loops = TRUE, normalized = TRUE))[degree(g2, loops = TRUE, normalized = TRUE)==max(degree(g2, loops = TRUE, normalized = TRUE))]

#Again, the students with the max degrees are those who belong to the cluster.
```