---
title: "Assignment 2"
author: "Yi Chen"
date: "September 26, 2020"
output: html_document
---
# Part I

## Data Wrangling
In the hackathon a project was proposed to collect data from student video watching, a sample of this data is available in the file video-data.csv.

stid = student id
year = year student watched video
participation = whether or not the student opened the video
watch.time = how long the student watched the video for
confusion.points = how many times a student rewatched a section of a video
key,points = how many times a student skipped or increased the speed of a video

```{r message=FALSE, warning=FALSE}
#Install the 'tidyverse' package or if that does not work, install the 'dplyr' and 'tidyr' packages.

#Load the package(s) you just installed

library(tidyverse)
library(tidyr)
library(dplyr)

D1 <- read.csv("video-data.csv", header = TRUE)
str(D1)

#Create a data frame that only contains the years 2018
D2 <- filter(D1, year == 2018)
```

## Histograms
```{r}
#Generate a histogram of the watch time for the year 2018

hist(D2$watch.time)

#Change the number of breaks to 100, do you get the same impression?

hist(D2$watch.time, breaks = 100)

#Cut the y-axis off at 10

hist(D2$watch.time, breaks = 100, ylim = c(0,10))

#Restore the y-axis and change the breaks so that they are 0-5, 5-20, 20-25, 25-35

hist(D2$watch.time, breaks = c(0,5,20,25,35))

```

## Plots
```{r}
#Plot the number of confusion points against the watch time

plot(D1$confusion.points, D1$watch.time)

#Create two variables x & y
x <- c(1,3,2,7,6,4,4)
y <- c(2,4,2,3,2,4,3)

#Create a table from x & y
table1 <- table(x,y)

#Display the table as a Barplot
barplot(table1)

#Create a data frame of the average total key points for each year and plot the two against each other as a lines

D3 <- D1 %>% group_by(year) %>% summarise(mean_key = mean(key.points))

plot(D3$year, D3$mean_key, type = "l", lty = "dashed")

#Create a boxplot of total enrollment for three students
D4 <- filter(D1, stid == 4|stid == 20| stid == 22)
#The drop levels command will remove all the schools from the variable with no data  
D4 <- droplevels(D4)
boxplot(D4$watch.time~D4$stid, xlab = "Student", ylab = "Watch Time")
```
## Pairs
```{r}
#Use matrix notation to select columns 2, 5, 6, and 7
D5 <- D1[,c(2,5,6,7)]
#Draw a matrix of plots for every combination of variables
pairs(D5)
```

## Part II

1. Create a simulated data set containing 100 students, each with a score from 1-100 representing performance in an educational game. The scores should tend to cluster around 75. Also, each student should be given a classification that reflects one of four interest groups: sport, music, nature, literature.

```{r}
set.seed(3356)
simulated_data <- data.frame(
  "performance" = rnorm(n = 100,mean = 75,sd = 20),
  "group" = factor(sample(c("sport", "music", "nature", "literature"),100,replace = T))
)


simulated_data <- simulated_data %>%
  mutate(performance = pmax(performance,0)) %>%
  mutate(performance = pmin(performance,100)) %>%
  mutate(performance = round(performance,2)) %>%
  add_column(Sid = 1:100) %>%
  select(Sid,group,performance)

head(simulated_data)
```

2. Using base *R commands*, draw a histogram of the scores. Change the breaks in your histogram until you think they best represent your data.

```{r}
hist(simulated_data$performance,
     xlim = c(30,100),
     breaks = 7,
     main = "Histogram of Student Scores",
     xlab = "Score",
     ylab = "Density")
```


3. Create a new variable that groups the scores according to the breaks in your histogram.

```{r}
simulated_data <- simulated_data %>% mutate(
  score_group =cut(simulated_data$performance,breaks = 7,labels = letters[1:7])
)
head(simulated_data)

```

4. Now using the colorbrewer package (RColorBrewer; http://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3) design a pallette and assign it to the groups in your data on the histogram.

```{r}
library(RColorBrewer)
#display.brewer.all()
#display.brewer.all(colorblindFriendly = TRUE)
#display.brewer.pal(n = 7, name = 'Dark2')

hist(simulated_data$performance,
     xlim = c(30,100),
     breaks = 7,
     main = "Histogram of Student Scores",
     xlab = "Score",
     ylab = "Density",
     col =brewer.pal(n = 7, name = 'Dark2'))
```


5. Create a boxplot that visualizes the scores for each interest group and color each interest group a different color.

```{r}
#Make a vector of the colors from RColorBrewer
  boxplot(simulated_data$performance~simulated_data$score_group, 
          xlab = "Score Level", 
          ylab = "Performance (Score)",
          main = "boxplot of score over levels",
          col = brewer.pal(n = 7, name = 'Dark2'))
```


6. Now simulate a new variable that describes the number of logins that students made to the educational game. They should vary from 1-25.

```{r}
simulated_data <- simulated_data %>%
  mutate(logins = sample(1:25,100,replace = T))
head(simulated_data)
```

7. Plot the relationships between logins and scores. Give the plot a title and color the dots according to interest group.

```{r}
plot(simulated_data$performance, simulated_data$logins,type = "p",
     ylab = "logins",xlab = "Performance (Scores)",
     main = "Scatter Plot of Logins and Scores")

```


8. R contains several inbuilt data sets, one of these in called AirPassengers. Plot a line graph of the the airline passengers over time using this data set.

```{r}
plot(AirPassengers, type = "l")
```


9. Using another inbuilt data set, iris, plot the relationships between all of the variables in the data set. Which of these relationships is it appropraiet to run a correlation on? 

```{r}
pairs(iris)
str(iris)
pairs(iris %>% select(-Species))
```

*Answer*: Pearson correlation is designed for the continuous variables. In iris dataset, species is a categoriacal variable, which is not suitable for correaltion. All the other variables are fine.


# Part III - Analyzing Swirl

## Data 

In this repository you will find data describing Swirl activity from the class so far this semester. Please connect RStudio to this repository.

### Instructions
  
1. Insert a new code block
2. Create a data frame from the `swirl-data.csv` file called `DF1`
```{r}
DF1 <- read.csv("swirl-data.csv", header = TRUE)
str(DF1)
head(DF1)
```


The variables are:

`course_name` - the name of the R course the student attempted  
`lesson_name` - the lesson name  
`question_number` - the question number attempted
`correct` - whether the question was answered correctly  
`attempt` - how many times the student attempted the question  
`skipped` - whether the student skipped the question  
`datetime` - the date and time the student attempted the question  
`hash` - anonymyzed student ID  

3. Create a new data frame that only includes the variables `hash`, `lesson_name` and `attempt` called `DF2`
```{r}
DF2 <- DF1 %>%
  select(hash,lesson_name,attempt)
head(DF2)
```


4. Use the `group_by` function to create a data frame that sums all the attempts for each `hash` by each `lesson_name` called `DF3`
```{r}
DF3 <- DF2 %>%
  group_by(hash,lesson_name) %>%
  summarise(attemp = sum(attempt))
head(DF3)
```


5. On a scrap piece of paper draw what you think `DF3` would look like if all the lesson names were column names

| harsh | lesson_name_1 | lesson_name_2 | lesson_name_3 | lesson_name_4 |
| :---: | :-----------: | :-----------: | :-----------: | :-----------: |
| 2864  | 29            | 49            | 51            | 116           |

6. Convert `DF3` to this format  
```{r}
unique(DF3$lesson_name)
## there is one lesson name missing
DF3 %>% filter(lesson_name=="") %>%
  distinct(attemp)
## there is no real information for lesson name as ""
## consequently, I will drop these records
DF3 <- DF3 %>% drop_na()

DF3_wide <- DF3 %>%
  pivot_wider(names_from = lesson_name,values_from = attemp)
head(DF3)

```


7. Create a new data frame from `DF1` called `DF4` that only includes the variables `hash`, `lesson_name` and `correct`
```{r}
DF4 <- DF1 %>%
  select(hash,lesson_name,correct)
head(DF4)
```


8. Convert the `correct` variable so that `TRUE` is coded as the **number** `1` and `FALSE` is coded as `0`  
```{r}
DF4 <- DF4 %>%
  mutate(correct = ifelse(correct=="TRUE",1,0))
head(DF4)
```


9. Create a new data frame called `DF5` that provides a mean score for each student on each course
```{r}
# for missing correct, I will score then as false
# the mean score is the correct of attempts
DF5 <- DF1 %>%
  replace_na(list(correct="FALSE")) %>%
  mutate(correct = ifelse(correct=="TRUE",1,0)) %>%
  group_by(hash,lesson_name) %>%
  summarise(average_score = round(mean(correct) * 100,2)) %>%
  arrange(desc(average_score))
head(DF5)
```


10. **Extra credit** Convert the `datetime` variable into month-day-year format and create a new data frame (`DF6`) that shows the average correct for each day
```{r}
library(lubridate)
DF6 <- DF1 %>%
  drop_na() %>%
  mutate(datetime =
          format(
            as.POSIXct(datetime, origin = '1970-01-01', tz = 'GMT'),"%m-%d-%Y")) %>%
  replace_na(list(correct="FALSE")) %>%
  mutate(correct = ifelse(correct=="TRUE",1,0)) %>%
  group_by(datetime) %>%
  summarise(average_correct = round(mean(correct),2)) %>%
  arrange(datetime)
head(DF6)


## note: there are some very small time-steamps, which may be error.
```



Finally use the knitr function to generate an html document from your work. Commit, Push and Pull Request your work back to the main branch of the repository. Make sure you include both the .Rmd file and the .html file. 
