---
title: "Assignment2"
author: "Cy"
date: "10/1/2020"
output: html_document
---

# Part 1

## Import data
```{r}
#Install the 'tidyverse' package or if that does not work, install the 'dplyr' and 'tidyr' packages.
#Load the package(s) you just installed
library(tidyr)
D1 <- read.csv("video-data.csv", header = TRUE)
#Create a data frame that only contains the years 2018
D2 <- filter(D1, year == 2018)
```

## Histograms
```{r}
#Generate a histogram of the watch time for the year 2018
hist(D2$watch.time)
#Change the number of breaks to 100, do you get the same impression?
hist(D2$watch.time, breaks = 100)
#Cut the y-axis off at 10
hist(D2$watch.time, breaks = 100, ylim = c(0,10))
#Restore the y-axis and change the breaks so that they are 0-5, 5-20, 20-25, 25-35
hist(D2$watch.time, breaks = c(0,5,20,25,35))

```
## Plots
```{r}
#Plot the number of confusion points against the watch time
plot(D1$confusion.points, D1$watch.time)
#Create two variables x & y
x <- c(1,3,2,7,6,4,4)
y <- c(2,4,2,3,2,4,3)
#Create a table from x & y
table1 <- table(x,y)
#Display the table as a Barplot
barplot(table1)
#Create a data frame of the average total key points for each year and plot the two against each other as a lines
D3 <- D1 %>% group_by(year) %>% summarise(mean_key = mean(key.points))
plot(D3$year, D3$mean_key, type = "l", lty = "dashed")
#Create a boxplot of total enrollment for three students
D4 <- filter(D1, stid == 4|stid == 20| stid == 22)
#The drop levels command will remove all the schools from the variable with no data  
D4 <- droplevels(D4)
boxplot(D4$watch.time~D4$stid, xlab = "Student", ylab = "Watch Time")
```

## Pairs
```{r}
#Use matrix notation to select columns 2, 5, 6, and 7
D5 <- D1[,c(2,5,6,7)]
#Draw a matrix of plots for every combination of variables
pairs(D5)
#Correlations between variables
cor (D5)
```


# Part 2

## Create a simulated data set
```{r}
S1a <- function(n, m, s, lwr, upr, nnorm) {
set.seed(1)
samp <- rnorm(nnorm, m, s)
samp <- samp[samp >= lwr & samp <= upr]
if (length(samp) >= n) {
return(sample(samp, n))
}  
stop(simpleError("Not enough values to sample from. Try increasing nnorm."))
} 
scores <- S1a(n=100, m=75, s=10, lwr=50, upr=100, nnorm=1000000)
round(scores)
interest <- c("sport", "music", "nature", "literature")
S1 <- data.frame(scores)
S1$interest <-sample(interest, 100, replace = TRUE)
summary(S1)

```

## draw a histogram of the scores
```{r}
hist(S1$scores, breaks = 10，xlim=c(50,100))
```

## Create a new variable that groups the scores according to the breaks in your histogram
```{r}
label1 <- letters [1:10]
S1$breaks <- cut(S1$scores, breaks = 10, labels = label1)
```

## design a pallette and assign it to the groups in your data on the histogram
```{r}
library (RColorBrewer)
display.brewer.all()
S1$colors <- brewer.pal(10, "Set3")
hist(S1$scores, col = S1$colors)
```

## Create a boxplot that visualizes the scores for each interest group and color each interest group a different color
```{r}
interest.col <- brewer.pal(4, "Pastel1")
boxplot(S1$scores~S1$interest, col = interest.col, xlab = "Student Interest Group", ylab = "Student Scores")
```

## simulate a new variable that describes the number of logins that students made to the educational game. They should vary from 1-25.
```{r}
logins <- seq(from=1, to=25)
S1$logins <-sample(logins, 100, replace = TRUE)
```

## Plot the relationships between logins and scores. Give the plot a title and color the dots according to interest group.
```{r}
plot(S1$scores, S1$logins, col = interest.col, ylab = "Student Logins", xlab = "Student Scores")
```

## Plot a line graph of the the airline passengers over time using the AirPassengers data set.
```{r}
plot(AirPassengers)
```

## Using another inbuilt data set, iris, plot the relationships between all of the variables in the data set. Which of these relationships is it appropraiet to run a correlation on? 
```{r}
pairs(iris)
cor(iris$Sepal.Length,iris$Petal.Length)
cor(iris$Petal.Length,iris$Petal.Width)
cor(iris$Sepal.Length,iris$Sepal.Width)
```

# Part 3

## Import data
```{r}
DF1 <- read.csv("swirl-data.csv", header = TRUE)
```

## Create a new data frame that only includes the variables `hash`, `lesson_name` and `attempt` called `DF2`
```{r}
myvars <- c("hash", "lesson_name", "attempt")
DF2 <- DF1[myvars]
```

