# Assignment 2 - Social Network Analysis

## Part I& II
Start by installing the "igraph" package. Once you have installed igraph, load the package.

Now upload the data file "discipline-data.csv" as a data frame called "D1". Each row is a disciplinary action from a teacher to a student so the first line shows that teacher "E" sent student "21" to the principal. It also shows the gender of both the teacher and student and the student's main elective field of study ("major"") and the field that the teacher instructs in ("t.expertise").

Before you proceed, you will need to change the data type of the student id variable. Since it is a number R will automatically think it is an integer and code it as such (look at the list of variables by clicking on the data frame arrow in the Data pane. Here you will see the letters "int"" next to the stid variable, that stands for integer). However, in this case we are treating the variable as a category, there is no numeric meaning in the variable. So we need to change the format to be a category, what R calls a "factor". We can do this with the following code:

```{r}
#load packages
library(igraph)
library(dplyr)
library(tidyr)
library(readr)
library(base)
#install data
D1 <- read_csv("discipline-data.csv")
D1$stid <- as.factor(D1$stid)
#isolate variables within D2
D2 <- dplyr::select(D1, tid, stid)
#collapse into single row and show how many times teacher-student pairs appear
EDGE <- dplyr::count(D2, tid, stid)
names(EDGE) <- c("from", "to", "count")
#make the vertex list of all the teachers and students and their characteristics in our network
#seperate teachers
V.TCH <- dplyr::select(D1, tid, t.gender, t.expertise)
#Remove all the repeats so that we just have a list of each teacher and their characteristics
V.TCH <- unique(V.TCH)
#Add a variable that describes that they are teachers
V.TCH$group <- "teacher"
#Now repeat this process for the students
V.STD <- dplyr::select(D1, stid, s.gender, s.major)
V.STD <- unique(V.STD)
V.STD$group <- "student"
#Make sure that the student and teacher data frames have the same variables names
names(V.TCH) <- c("id", "gender", "topic", "group")
names(V.STD) <- c("id", "gender", "topic", "group")
#Bind the two data frames together (you will get a warning because the teacher data frame has 5 types of id (A,B,C,D,E) and the student has 25 (1-30), this isn't a problem)
VERTEX <- dplyr::bind_rows(V.TCH, V.STD)
```
```{r}
#First we will make an object that contains the graph information using our two dataframes EDGE and VERTEX. Notice that we have made "directed = TRUE" - our graph is directed since discipline is being given from a teacher to a student.
g <- graph.data.frame(EDGE, directed=TRUE, vertices=VERTEX)
#Now we can plot our graph using the force directed graphing technique - our old friend Fruchertman-Reingold!
plot(g,layout=layout.fruchterman.reingold)
#There are many ways to change the attributes of the graph to represent different characteristics of the newtork. For example, we can color the nodes according to gender.
gender_color <- ifelse(VERTEX$gender == "female", "lightblue", "orange")
plot(g,layout=layout.fruchterman.reingold, vertex.color=gender_color)
#We can change the thickness of the edge according to the number of times a particular teacher has sent a particular student to the principal.
plot(g,layout=layout.fruchterman.reingold, vertex.color=gender_color, edge.width=EDGE$count)
V(g)$vertex_degree <-  degree(g)
#change verticies size based on student vertices in terms of the number of disciplinary actions they have recieved
plot(g,layout=layout.fruchterman.reingold, vertex.color=gender_color, edge.width=EDGE$count,
     vertex.size=V(g)$vertex_degree)
```
##Part III
###Data Cleaning 
```{r}
#import class data
D3 <- read_csv("HUDK4050-2018-classes.csv")
#filter relevant variables
D3 <- unite(D3, "Name", c(Q8, Q9), sep=" ")
D4 <- dplyr::select(D3, Name, Q1, Q3, Q4, Q5, Q6, Q7)
D4 <- D4[-c(1, 2, 3, 39), ]
#fix rows with no space between dept and course number
D4[3,3] = "HUDK 5035"
D4[3,4] = "HUDK 4011"
D4[3,5] = "HUDM 4125"
D4[3,2] = "HUDK 4050"
D4[4,3] = "MSTU 5003"
D4[4,4] = "MSTU 4083"
D4[4,5] = "MSTU 4039"
D4[4,6] = "MSTU 4000"
D4[36,2] = "HUDK 4050"
D4[36,3] = "EDPA 6002"
#gather
D4G <- gather(D4, Class, Count, -1)
colnames(D4G) <- c("Name", "Count", "Class")
D4G$Count <- 1
#spread so class is in column with 0 or 1 indicating if student is enrolled
#change NA to 0  
D4G <- na.omit(D4G)
#change string to 0s 
D5 <- spread(D4G, Class, Count)
#remove second and last columns
D5[-1] <- lapply(D5[-1], function(x) as.integer(x!="0"))
D5[is.na(D5)] <- 0 
D5 <- D5[,-19]
```

###Clustering and Social Network Analysis
```{r}
library(klaR)
#cluster
fit <- kmodes(D5, 4, iter.max = 10, weighted = FALSE)
#change names to index
row.names(D5) <- D5$Name
#remove name column 
D5$Name <- NULL
#convert to matrix
m1 <- as.matrix(D5)
m2 <- m1 %*% t(m1)
diag(m2) <-0
plot(graph_from_adjacency_matrix(m2, mode="undirected", weighted=NULL, diag =TRUE), layout=layout.fruchterman.reingold, vertex.color=fit$cluster)
```
