---
title: "Assignment 6 - Social Network Analysis"
author: "Charles Lang"
date: "Oct 13, 2016"
output: html_document
---

In Assignment 2 we will be looking at some disciplinary data from a middle school. The file "discipline-data.csv" shows which teacher sent which student to the principal during the semester.

We will be using the "igraph" package to visualize the disciplinary relationships between teachers and students as a network. You can read more about igraph here:

http://igraph.org/r/

Start by installing the "igraph" package. Once you have installed igraph, load the package.

Now upload the data file "discipline-data.csv" as a data frame called "D1". Each row is a disciplinary action from a teacher to a student so the first line shows that teacher "E" sent student "21" to the principal. It also shows the gender of both the teacher and student and the student's main elective field of study ("major"") and the field that the teacher instructs in ("t.expertise").

Before you proceed, you will need to change the data type of the student id variable. Since it is a number R will automatically think it is an integer and code it as such (look at the list of variables by clicking on the data frame arrow in the Data pane. Here you will see the letters "int"" next to the stid variable, that stands for integer). However, in this case we are treating the variable as a category, there is no numeric meaning in the variable. So we need to change the format to be a category, what R calls a "factor". We can do this with the following code:

```{r}
D1$stid <- as.factor(D1$stid)
```

igraph requires data to be in a particular structure. There are several structures that it can use but we will be using a combination of an "edge list" and a "vertex list". As you might imagine the edge list contains a list of all the relationships between students and teachers and any characteristics of those edges that we might be interested in. There are two essential variables in the edge list a "from" variable and a "to" variable that descibe the relationships between vertices (a disciplinary action is given "from" and teacher "to" a student). While the vertix list contains all the characteristics of those vertices, in our case gender and major.

So let's convert our data into an edge list!

First we will isolate the variables that are of interest: tid and stid

```{r}
library(dplyr)

D2 <- dplyr::select(D1, tid, stid)
```

Since our data represnts every time a teacher sends a student to the principal there are multiple rows when the same teacher sends the same student. We want to collapse these into a single row, with a variable that shows how many times a teacher-student pair appears.

```{r}

EDGE <- dplyr::count(D2, tid, stid)

names(EDGE) <- c("from", "to", "count")

```

EDGE is your edge list. Now we need to make the vertex list, a list of all the teachers and students and their characteristics in our network.

```{r}

#First we will separate the teachers from our original data frame
V.TCH <- dplyr::select(D1, tid, t.gender, t.expertise)
#Remove all the repeats so that we just have a list of each teacher and their characteristics
V.TCH <- unique(V.TCH)
#Add a variable that describes that they are teachers
V.TCH$group <- "teacher"

#Now repeat this process for the students
V.STD <- dplyr::select(D1, stid, s.gender, s.major)
V.STD <- unique(V.STD)
V.STD$group <- "student"

#Make sure that the student and teacher data frames have the same variables names
names(V.TCH) <- c("id", "gender", "topic", "group")
names(V.STD) <- c("id", "gender", "topic", "group")

#Bind the two data frames together (you will get a warning because the teacher data frame has 5 types of id (A,B,C,D,E) and the student has 25 (1-30), this isn't a problem)
VERTEX <- dplyr::bind_rows(V.TCH, V.STD)

```

Now we have both a Vertex and Edge list it is time to plot our graph!

```{r}
#Load the igraph package

library(igraph)

#First we will make an object that contains the graph information using our two dataframes EDGE and VERTEX. Notice that we have made "directed = TRUE" - our graph is directed since discipline is being given from a teacher to a student.

g <- graph.data.frame(EDGE, directed=TRUE, vertices=VERTEX)

#Now we can plot our graph using the force directed graphing technique - our old friend Fruchertman-Reingold!

plot(g,layout=layout.fruchterman.reingold)

#There are many ways to change the attributes of the graph to represent different characteristics of the newtork. For example, we can color the nodes according to gender.

plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender)

#We can change the thickness of the edge according to the number of times a particular teacher has sent a particular student to the principal.

plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender, edge.width=EDGE$count)

#Other attributes you may want to change are:

#Arrow size
edge.arrow.size=

#Remove the labels
vertex.label=NA
```

```

#Vertex size
vertex.size=
  
#And many others!
```

Play around with different graphs and then save your final version to the repo. Create a Zotero item for this R program and then commit, push and pull.

Once you have done this you need to practice with data from our class. Please create a **person-network** with the data set hudk4050-classes.csv. Once you have done this, also [look up](http://igraph.org/r/) how to generate the following network metrics: betweeness centrality and dregree. **Who is the most central person in the network?**

```{r}
#make a new column for combined name
classes$name <- do.call(paste, c(classes[c("First.name", "Last.name")], sep = " ")) 
#take out first and last name columns by creating new table
 C2 <- dplyr::select(classes, name, Course1, Course2, Course3, Course4, Course5)
 #make all the courses into one long column
 C3 <- tidyr::gather(C2, "course", col = 2:6, na.rm = TRUE)
C4 <- tidyr::separate(C3, "name", "value") 
#remove blanks from value (list of courses)
C4 <- tidy::filter(C3, "value" > 0 )
#take out list of course1, etc.
 C4 <- dplyr::select(C4, name, value)
#add in column for count 
> C4$count <- "1"
 #change table for courses as columns with count for individuals
> C5 <- tidyr::spread(C4, "value", "count")
 #turn NAs into 0s
  C5[is.na(C5)] <- 0
  
  
  #trying to add count instead of "1"
  C4 <- dplyr::count(C4, name, value)
  #remove rows where no course is listed
  C4 <- dplyr::filter(C4, value > 0)
  #reshape
  C5 <- tidyr::spread(C4, "value", "n")
  
  C6 <- as.matrix(C5)
  #non numeric error. 
  #remove NAs
  C6[is.na(C6)] <- 0
  #still non numeric

  
  
  > colnames(C6) <- as.character(C6[1,])
> View(C6)
> C6 <- dplyr::filter(C6, )
  
  
  
  
  ###### Starting over. 
  
  #make the new full name column
  E$Name <- do.call(paste, c(E[c("First.name", "Last.name")], sep = " "))
  
  #new data frame with Name and course 1-5
  E1 <- dplyr::select(E, Name, Course1, Course2, Course3, Course4, Course5)
  
  #combine all courses into one column
  E2 <- tidyr::gather(E1, "number", "Course", col = 2:6, na.rm = TRUE)
  
  #drop course number column
  E2 <- dplyr::select(E2, Name, Course)
  
  #remove rows with no course listed
  E2 <- dplyr::filter(E2, Course > 0)
  
  #add count 
  E3 <- dplyr::count(E2, Name, Course)
  
  #make courses into columns with 0 or 1 against names
  E4 <- tidyr::spread(E3, Course, n, fill = 0)
  
  #now the problem: name is a column with names (characters) whereas I want it to be the row names so that all the table data is 1 or 0
  
  #try with Name as columns instead of course
  E5 <- tidyr::spread(E3, Name, n, fill = 0)
  
  #remove columb with Course names
  E6 <- dplyr::select(E5, -Course)
  
  #set people (and an extra, because I don't know what diag might do yet *just removes connection with themselves)
  people <- E6 %% t(E6)
  people1 <- people
  
  #with diag
  diag(people1) <- NA
  
   #errors. quitting fot the day.
  
  
  #restarting fresh, again. 
  
  #set up initial table view
  hud <- read.csv("hudk4050-classes.csv")
  #View(hud)
  
  #create column for full name
  hud$Name <- paste(hud$First.name, hud$Last.name, sep = " ")
  
  #new table with names and five columns of course names
  hud1 <- dplyr::select(hud, Name, Course1, Course2, Course3, Course4, Course5)
  
  #new table with all courses in one column
   hud2 <- tidyr::gather(hud1, Number, Course, col = 2:6, na.rm = TRUE)
   
   #remove list of course number for person
   hud2 <- dplyr::select(hud2, Name, Course)
   
   #add count for people/courses in new table
   hud3 <- dplyr::count(hud2, Name, Course)
   
   #remove rows where course is empty 
   hud3 <- dplyr::filter(hud3, Course > 0)
   
  #plotted like student discipline above, not pretty but functional:
   h <- graph.data.frame(hud3, directed = FALSE, vertices = NULL)
   plot(h, layout=layout.fruchterman.reingold)
   
   #with direction, but not terribly meaningful here:
   h1 <- graph.data.frame(hud3, directed = TRUE, vertices = NULL)
   plot(h1, layout=layout.fruchterman.reingold)
   
   #stil error for revere style:
   people <- hud3 %% t(hud3)
   #Error in FUN(left, right) : non-numeric argument to binary operator
   
   people <- h %% t(h)
   #Error in h%%t(h) : non-numeric argument to binary operator

   ```
   
   
   
  
