# Assignment 2 - Social Network Analysis

## Part I
Start by installing the "igraph" package. Once you have installed igraph, load the package.

Now upload the data file "discipline-data.csv" as a data frame called "D1". Each row is a disciplinary action from a teacher to a student so the first line shows that teacher "E" sent student "21" to the principal. It also shows the gender of both the teacher and student and the student's main elective field of study ("major"") and the field that the teacher instructs in ("t.expertise").

Before you proceed, you will need to change the data type of the student id variable. Since it is a number R will automatically think it is an integer and code it as such (look at the list of variables by clicking on the data frame arrow in the Data pane. Here you will see the letters "int"" next to the stid variable, that stands for integer). However, in this case we are treating the variable as a category, there is no numeric meaning in the variable. So we need to change the format to be a category, what R calls a "factor". We can do this with the following code:

```{r}

D1 = read.csv("discipline-data.csv")
D1$stid <- as.factor(D1$stid)
```

igraph requires data to be in a particular structure. There are several structures that it can use but we will be using a combination of an "edge list" and a "vertex list". As you might imagine the edge list contains a list of all the relationships between students and teachers and any characteristics of those edges that we might be interested in. There are two essential variables in the edge list a "from" variable and a "to" variable that descibe the relationships between vertices (a disciplinary action is given "from" and teacher "to" a student). While the vertix list contains all the characteristics of those vertices, in our case gender and major.

So let's convert our data into an edge list!

First we will isolate the variables that are of interest: tid and stid

```{r}
library(dplyr)
library(tidyr)

D2 <- dplyr::select(D1, tid, stid)
```

Since our data represnts every time a teacher sends a student to the principal there are multiple rows when the same teacher sends the same student. We want to collapse these into a single row, with a variable that shows how many times a teacher-student pair appears.

```{r}

EDGE <- dplyr::count(D2, tid, stid)

names(EDGE) <- c("from", "to", "count")

```

EDGE is your edge list. Now we need to make the vertex list, a list of all the teachers and students and their characteristics in our network.

```{r}

#First we will separate the teachers from our original data frame
V.TCH <- dplyr::select(D1, tid, t.gender, t.expertise)
#Remove all the repeats so that we just have a list of each teacher and their characteristics
V.TCH <- unique(V.TCH)
#Add a variable that describes that they are teachers
V.TCH$group <- "teacher"

#Now repeat this process for the students
V.STD <- dplyr::select(D1, stid, s.gender, s.major)
V.STD <- unique(V.STD)
V.STD$group <- "student"

#Make sure that the student and teacher data frames have the same variables names
names(V.TCH) <- c("id", "gender", "topic", "group")
names(V.STD) <- c("id", "gender", "topic", "group")

#Bind the two data frames together (you will get a warning because the teacher data frame has 5 types of id (A,B,C,D,E) and the student has 25 (1-30), this isn't a problem)
VERTEX <- dplyr::bind_rows(V.TCH, V.STD)

```

Now we have both a Vertex and Edge list it is time to plot our graph!

```{r}
#Load the igraph package

library(igraph)

#First we will make an object that contains the graph information using our two dataframes EDGE and VERTEX. Notice that we have made "directed = TRUE" - our graph is directed since discipline is being given from a teacher to a student.

g <- graph.data.frame(EDGE, directed=TRUE, vertices=VERTEX)

#Now we can plot our graph using the force directed graphing technique - our old friend Fruchertman-Reingold!

plot(g,layout=layout.fruchterman.reingold)

#There are many ways to change the attributes of the graph to represent different characteristics of the newtork. For example, we can color the nodes according to gender.

plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender)

#We can change the thickness of the edge according to the number of times a particular teacher has sent a particular student to the principal.

plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender, edge.width=EDGE$count)
```

## Part II

In Part II your task is to [look up](http://igraph.org/r/) in the igraph documentation and create a graph that sizes the student vertices in terms of the number of disciplinary actions they have recieved, and the teachers in terms of the number of disciplinary actions they have given out.


```{r}
# student table
student_data <- select(EDGE,to,count)
# sum up incidents per student
student_data <- group_by(student_data,to)
student_data <- summarise(student_data, count=sum(count))
student_data <- rename(student_data,id=to)
# teacher table
teacher_data <- select(EDGE,from,count)
# sum up incidents per teacher
teacher_data <- group_by(teacher_data,from)
teacher_data <- summarise(teacher_data, count=sum(count))
teacher_data <- rename(teacher_data,id=from)
# merge student/teacher data
full_data <- full_join(student_data,teacher_data,by="id")
full_data <- group_by(full_data,id)
full_data <- transmute(full_data,count=sum(count.x,count.y,na.rm=TRUE))
# merge into vertex data
VERTEX <- left_join(VERTEX,full_data,by="id")
# size vertices in proportion to number of incidences (# * 3.5 for visibility)
plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender, vertex.size=(VERTEX$count*3.5))
```

## Part III

Now practice with data from our class. Please create a **person-network** with the data set hudk4050-classes.csv. To create this network you will need to create a person-class matrix using the tidyr functions and then create a person-person matrix using `t()`. You will then need to plot a matrix rather than a data frame using igraph.

Once you have done this, also [look up](http://igraph.org/r/) how to generate the following network metrics: betweeness centrality and dregree. **Who is the most central person in the network?**
```{r}
# formats non-empty class names
# if class name contains a space ("HUDK 4050"), leave it alone; if not, then add a space where appropriate
format_class <- function(e) {
  s <- as.character(e)
  if((nchar(s)>0) && (grepl(" ",s,fixed=TRUE)==1)) { # string contains space
      s
  } else { # no space
      paste(substring(s,1,4),substring(s,5),sep=" ")
  }
}

class_data <- read.csv("HUDK4050-2018-classes.csv")
# get rid of headers, and also blank/duplicate data
class_data <- class_data[-c(1,2,3,39),]
# subset down to name and class columns
class_data <- select(class_data, Q8,Q9,Q1,Q3,Q4,Q5,Q6)

# rename columns for clarity
class_data <- rename(class_data,first_name=Q8,last_name=Q9,class1=Q1,class2=Q3,class3=Q4,class4=Q5,class5=Q6)
# also reformat data in dataframe
# combine first name + last name into single name
class_data <- transmute(class_data, name=paste(first_name,last_name,sep=" "),class1,class2,class3,class4,class5)
# mostly done cleaning data, aside from something to apply later

# construct person-class table
# gather classes into 1 column, adding (and then removing) a filler column
# there's probably a neater way to do this, but I can't figure it out...
class_data <- gather(class_data,key="discard_me",value="class",2:6)
class_data <- select(class_data,name,class)
# finish formatting all classes, in particular deleting blanks and adding spaces to things like "HUDK4050"
class_data <- class_data[(class_data$class != ""),]
class_data$class <- sapply(class_data$class,format_class)
# add a "taken" column of all 1's (indicating each student is taking their associated class(es))
class_data$taken <- 1
# spread the table out into a matrix: 1 if student is taking a given class, NA otherwise
class_wdata <- spread(class_data,class,taken)
# change all NAs to 0, so the matrix is full of 1s/0s for taking/not taking
class_wdata[is.na(class_wdata)] <- 0

# create the person-class matrix, excluding the name fields to make a numeric matrix
pc_mtx <- select(class_wdata,-name)
pc_mtx <- as.matrix(pc_mtx)

# create the person-person matrix by multiplying the person-class matrix by its transposition
pp_mtx <- pc_mtx %*% t(pc_mtx)
# add row/column names, i.e., student names
rownames(pp_mtx) <- class_wdata$name
colnames(pp_mtx) <- class_wdata$name
# result: pp_mtx contains the number of classes each person shares with each other person (1-5, since everyone is taking HUDK 4050) => undirected adjacency matrix weighted by number of shared classes

# create graph from this person-person adjacency matrix; ignore diagonals, as we're not interested in how many classes a person shares with themselves
g2 <- graph_from_adjacency_matrix(pp_mtx, mode = c("undirected"), weighted = TRUE, diag = FALSE,add.colnames = NULL, add.rownames = NA)
#plot the graph
plot(g2,layout=layout.fruchterman.reingold,vertex.label.cex=0.5,width=10)

# create a "sparse" graph that considers shared classes OUTSIDE of HUDK 4050
pp_mtx_sparse <- pp_mtx - 1 # just subtract 1 from all elements, which will discount the class everyone shares
g3 <- graph_from_adjacency_matrix(pp_mtx_sparse, mode = c("undirected"), weighted = TRUE, diag = FALSE,add.colnames = NULL, add.rownames = NA)
# plot this as well
plot(g3,layout=layout.fruchterman.reingold,vertex.label.cex=0.5,width=10)

# Part 3.5: looking for the most central person(s)
#betweenness centrality - look for the students with least score for the most central
# use the original "dense" graph, since we want to consider all paths (shared classes)
bc <- betweenness(g2)
most_btwn_students <- names(bc[bc==min(bc)]) # get names of students with minimum score
# degree - look for the students highest degrees
# use the "sparse" graph, since the dense graph gives everyone an equal degree of 38 due to sharing HUDK 4050
dc <- degree(g3)
most_connected_students <- names(dc[dc==max(dc)])

# create a vector of central students by combining the above results and removing any duplicates
most_central_students = unique(c(most_btwn_students,most_connected_students))
```
