---
title: "He Zhang-Assignment 2"
author: "He Zhang"
date: "10/19/2017"
output: html_document
---

#Part 1
#install igraph package
```{r}
# Download and install the package
install.packages("igraph")

# Load package
library(igraph)
 
```

#upload data file "discipline-data.csv" as a data frame called "D1"
```{r}
D1<-read.csv("~/Desktop/He Zhang-Assignment 2/discipline-data.csv", header=T)
D1
```

#change the format of student id variable from number to factor
```{r}
D1$stid <- as.factor(D1$stid)
```

#Isolate the variables that are of interest: tid and stid
```{r}
library(dplyr)
D2 <- dplyr::select(D1, tid, stid)
```

#frequency of a teacher-student pair
```{r}
EDGE <- dplyr::count(D2, tid, stid)
names(EDGE) <- c("from", "to", "count")
```

```{r}

#First we will separate the teachers from our original data frame
V.TCH <- dplyr::select(D1, tid, t.gender, t.expertise)
#Remove all the repeats so that we just have a list of each teacher and their characteristics
V.TCH <- unique(V.TCH)
#Add a variable that describes that they are teachers
V.TCH$group <- "teacher"

#Now repeat this process for the students
V.STD <- dplyr::select(D1, stid, s.gender, s.major)
V.STD <- unique(V.STD)
V.STD$group <- "student"

#Make sure that the student and teacher data frames have the same variables names
names(V.TCH) <- c("id", "gender", "topic", "group")
names(V.STD) <- c("id", "gender", "topic", "group")

#Bind the two data frames together (you will get a warning because the teacher data frame has 5 types of id (A,B,C,D,E) and the student has 25 (1-30), this isn't a problem)
VERTEX <- dplyr::bind_rows(V.TCH, V.STD)

```

```{r}
#Load the igraph package

library(igraph)

#First we will make an object that contains the graph information using our two dataframes EDGE and VERTEX. Notice that we have made "directed = TRUE" - our graph is directed since discipline is being given from a teacher to a student.

g <- graph.data.frame(EDGE, directed=TRUE, vertices=VERTEX)

#Now we can plot our graph using the force directed graphing technique - our old friend Fruchertman-Reingold!

plot(g,layout=layout.fruchterman.reingold)

#There are many ways to change the attributes of the graph to represent different characteristics of the newtork. For example, we can color the nodes according to gender.

plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender)

#We can change the thickness of the edge according to the number of times a particular teacher has sent a particular student to the principal.

plot(g,layout=layout.fruchterman.reingold, vertex.color=VERTEX$gender, edge.width=EDGE$count)

```

#Part 2
```{r}
library (dplyr)
#First we will separate the teachers from our original data frame
V.TCH2 <- dplyr::select(D1, tid, t.gender, t.expertise)
#Add a variable that describes that they are teachers
V.TCH2$group <- "teacher"

#Now repeat this process for the students
V.STD2 <- dplyr::select(D1, stid, s.gender, s.major)
V.STD2$group <- "student"

#Make sure that the student and teacher data frames have the same variables names
names(V.TCH2) <- c("id", "gender", "topic", "group")
names(V.STD2) <- c("id", "gender", "topic", "group")

#Add frequencies to V.TCH2 and V.STD2
library(data.table)
dt.STD2 = data.table(V.STD2)
dt.STD2[,'freq-id':=.N, by=id]
dt.TCH2 = data.table(V.TCH2)
dt.TCH2[,'freq-id':=.N, by=id]

#Remove all the repeats so that we just have a list of each teacher and their characteristics
V.STD2<-unique(as.data.frame(dt.STD2))
V.TCH2<-unique(as.data.frame(dt.TCH2))

#Bind the two data frames together
VERTEX2<-dplyr::bind_rows(V.TCH2, V.STD2)

#Load the igraph package
library(igraph)

#First we will make an object that contains the graph information using our two dataframes EDGE and VERTEX2.
g2 <- graph.data.frame(EDGE, directed=TRUE, vertices=VERTEX2)

#Set up color and size of the vertices
V(g2)$color<-ifelse(VERTEX2$group=="student", "lightblue", "orange")

#Plot
plot(g2, layout = layout.fruchterman.reingold,
       vertex.size = VERTEX2$`freq-id`,
       vertex.frame.color= NA,
       vertex.label.color = "black",
       vertex.label.family = "sans",
       edge.width=EDGE$count,  
       edge.color="gray")
```


#Part 3
```{r}
library(dplyr)
library(tidyr)
D3 <- read.csv("~/Desktop/He Zhang-Assignment 2/hudk4050-classes.csv", header=T)

D3$Name<-paste(D3$First.name, D3$Last.name)
D3["First.name"]=NULL
D3["Last.name"]=NULL
D3<-D3%>%
  select(Name, everything())

D3<-tidyr::gather(D3, "n", "course", 2:6)
D3$n<-NULL

D3$value<-1
D3 <- dplyr::filter(D3, course > 0, course != "HUDK4050")

D4<-unique(D3)

D4 <- tidyr::spread(D4, key="course",value="value", fill = NA, convert = FALSE, drop = TRUE)

D4<-D4[c(-2)]
row.names(D4)<-D4$Name
D4$Name <- NULL

D4[is.na(D4)]<-0

#person-class matrix
D4.matrix<-as.matrix(D4)
D4.matrix

#Person-Person matrix
D4.matrix2<-t(D4.matrix)

D4.matrix3 <- D4.matrix %*% D4.matrix2
diag(D4.matrix3) <- NA

#plot a matrix using igraph
library(igraph)
g3<-graph.adjacency(D4.matrix3, weighted=T,mode="undirected")
g3<-simplify(g3, remove.multiple = TRUE, remove.loops = TRUE)

set.seed(3952)

V(g3)$label.color <- rgb(0, 0, .2, .8)
V(g3)$frame.color <- NA
egam <- (log(E(g3)$weight)+.4) / max(log(E(g3)$weight)+.4)
E(g3)$color <- rgb(.5, .5, 0, egam)
layout1 <- layout.fruchterman.reingold(g3)
plot(g3, layout=layout1)

plot(g3, layout=layout.kamada.kawai)
```

#Part3-2 Lauren Romine, Joshua Coleman, Robert Jackson, Zhuqian Zhou and Jonathan Stelman are the most central persons in this network.
```{r}
betweenness(g3, v = V(g3), directed = TRUE, weights = NULL,
  nobigint = TRUE, normalized = FALSE)

degree(g3, v = V(g3), mode = c("all", "out", "in", "total"),
  loops = TRUE, normalized = FALSE)
```




