---
title: "MyAssignment2"
author: "Elliot Hu-Au"
date: "October 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

R Markdown for HUDK 4050: Assignment 2

Part II
```{r}


E1<-dplyr::select(EDGE,count)
student_count<- EDGE %>% group_by(EDGE$to) %>% summarize(sum = sum(count))
names(student_count)[1]<-"id"
new_count<- merge(VERTEX,student_count,by="id")

g <- graph.data.frame(EDGE, directed=TRUE, vertices=all_totals)

teach_count<- EDGE %>% group_by(from) %>% summarize(sum(count))
names(teach_count)[2]<-"sum"
names(teach_count)[1]<-"id"
new_teach_count<- merge (VERTEX,teach_count, by="id")

all_totals<-dplyr::bind_rows(new_teach_count,new_count)
nodeSize<- setNames(c(all_totals$sum),c(all_totals$id))


plot(g,layout=layout.fruchterman.reingold, vertex.color=all_totals$gender, edge.width=EDGE$count, vertex.size=all_totals$sum*3)
```

## Part III

Now practice with data from our class. Please create a **person-network** with the data set hudk4050-classes.csv. To create this network you will need to create a person-class matrix using the tidyr functions and then create a person-person matrix using `t()`. You will then need to plot a matrix rather than a data frame using igraph.

Once you have done this, also [look up](http://igraph.org/r/) how to generate the following network metrics: betweeness centrality and dregree. **Who is the most central person in the network?**


```{r}

class_2018<-read.csv("HUDK4050-2018-classes.csv")
class_2018<-select(class_2018,"Q8", "Q9", "Q1", "Q3", "Q4", "Q5", "Q6")
class_2018<- class_2018[-c(1,2,3,39),]
class_2018<-unite(class_2018, "name", Q8,Q9, sep = " ", remove= TRUE)

group_by(class_2018, Q1)
class_2018%>% summarise(name)

#creating the tall dataframe of names and classes
names_class1<-select(class_2018, name, Q1)

names_class2<-select(class_2018, name, Q3)
names_class2<-names_class2[-c(1,39),]

names_class3<-select(class_2018,name, Q4)
names_class3<-arrange(names_class3, Q4)
names_class3<-names_class3[-c(1:7),]

names_class4<-select(class_2018,name, Q5)
names_class4<-arrange(names_class4, Q5)
names_class4<-names_class4[-c(1:12),]

names_class5<-select(class_2018, name, Q6)
names_class5<-arrange(names_class5, Q6)
names_class5<-names_class5[-c(1:30),]

names(names_class1)<-c("name","class")
names(names_class2)<-c("name","class")
names(names_class3)<-c("name","class")
names(names_class4)<-c("name","class")
names(names_class5)<-c("name","class")

n_cl_total<-dplyr::bind_rows(names_class1,names_class2,names_class3, names_class4, names_class5)
n_cl_total<-arrange(n_cl_total, name)
n_cl_total$enroll<-1
#put a space in the class values where there is no space
n_cl_total$class<-gsub('([A-Z])([0-9])', '\\1 \\2', n_cl_total$class)


#create dataframe with each student as an observation and classes as each column
n2<-spread(n_cl_total,"class", enroll)
n2[is.na(n2)]<-0


#multiply matrix by transpose to get people to people matrix
n2_matrix<-as.matrix(n2, rownames.force = NA, dimnames(n2$name,class_list))
n3_matrix<-n2_matrix[,-1]
mode(n3_matrix)<-"numeric"
n3_tran<-n3_matrix %*% t(n3_matrix)

#giving names to the matrix dimensions
rownames(n3_tran)<-n2$name
colnames(n3_tran)<-n2$name
n4_tran<-n3_tran

#removing the HUDK4050 data since everyone is connected to each other through it
noHUDK4050mat<-n4_tran-1
diag(noHUDK4050mat)<-0
class_net<-graph.adjacency(noHUDK4050mat, mode="undirected", diag=FALSE)
plot(class_net, layout=layout_in_circle)

#betweenness calculation without HUDK4050 connections
B1<-betweenness(class_net, v = V(class_net), directed = FALSE, weights = NULL, normalized = FALSE)

#eigen centrality without HUDK 4050 connections
E1<-eigen_centrality(class_net, directed = FALSE, scale = TRUE, weights = NULL,
  options = arpack_defaults)

#degree without HUDK 4050 connections
Deg1<-degree(class_net, v = V(class_net), mode = c("all", "out", "in", "total"),loops = TRUE, normalized = FALSE)

```


Graph and calculations with HUDK 4050 connections included


```{r}

##graph with HUDK 4050 connections
n3_temp<- n3_tran
diag(n3_temp)<-0
class_net2<-graph.adjacency(n3_temp, mode="undirected", diag=FALSE)
plot(class_net2, layout=layout_in_circle)

#betweenness calculation WITH HUDK4050 connections
B2<-betweenness(class_net2, v = V(class_net2), directed = FALSE, weights = NULL, normalized = FALSE)

#eigen centrality WITH HUDK 4050 connections
E2<-eigen_centrality(class_net2, directed = FALSE, scale = TRUE, weights = NULL,
  options = arpack_defaults)

#degree wWITH HUDK 4050 connections
Deg2<-degree(class_net2, v = V(class_net2), mode = c("all", "out", "in", "total"),loops = TRUE, normalized = FALSE)




```



##Please ignore what is below. This was just test stuff.

```{r}
##stuff underneath this was test stuff...please ignore
#creating class list for
class_list<-unique(n_cl_total$class)
class_list<-as.data.frame(class_list)
class_list<-arrange(class_list, class_list)

#creating the EDGE data frame
class_2018_edge <- n_cl_total
C1<-n_cl_total
C1<- filter(n_cl_total,class == "HUDK 4050")

#arranged total students and classes alphabetically by class
C2<- arrange(n_cl_total, class)


#how many classes with more than 1 student
class_count<- as.data.frame(table(n_cl_total$class))
class_count <- subset(class_count,Freq != 1)

unique(n_distinct(n_cl_total$class))
names(class_2018_edge) <-(c("from", "to"))
```


